---
title: "Russian Housing Market"
output: pdf_document
---

```{r}
library(tidyverse)
library(caret)
library(DataExplorer)
library(rgdal)
library(RColorBrewer)
library(sp)
```

```{r}
train <- read.csv("train.csv")
test <- read.csv("test.csv")
head(train)

plot_missing(train[, 1:50])
plot_missing(train[, 51:100])
plot_missing(train[,101:150])
plot_missing(train[,151:200])
plot_missing(train[,201:250])
plot_missing(train[,251:292])

# Although these plots are easy, there are too many variables to have a single plot. As an alternative, I will calculate the percentage of missing values for each variable and then plot the ones that are greater than 0. 

missing_vals <- function(x) {
   value <-  round((sum(is.na(x))/length(x)) * 100, digits = 2) 
   return(value)
}
miss_pct <- map_dbl(train, missing_vals)

miss_pct.df <- data.frame(pct = miss_pct, Variable = names(miss_pct)) %>% 
    filter(pct > 0)

ggplot(miss_pct.df, aes(x = reorder(Variable, pct), y = pct)) + 
    geom_bar(stat = "identity", fill = "#DA291C") +
    labs(x = "Percentage of missing values", y = "Variable") +
    coord_flip()

nrow(miss_pct.df)
variable_type <- map_df(train, function(x){sum(is.numeric(x))})
names(variable_type)[variable_type == 0]
head(train$timestamp)
ncol(train)
```
There are a total of 292 columns in the training set. There are 41 variables that have more than 1% of values missing. There are 10 variables that have missing values, but 1% or less of the observations are missing that value. There are 276 numeric variables and 15 categorical variables and 1 variable that should be a date variable. As a result, I will change the timestamp to a date variable. 

I want to visualize the data on maps, so I'm going to start my EDA by doing that. I'm following a lot of this notebook (https://www.kaggle.com/jtremoureux/map-visualizations-with-external-shapefile), but hopefully will do some of my own stuff as well. 

```{r}
districts <- read_delim("districts.csv", delim = ";") %>% 
    select(ID, `English Name`, `Russian Name`)
areas <- read_delim("regions.csv", delim = ";")
shp <- readOGR(dsn = "moscow_districts/", layer = "moscow_adm")

# Get the box for polygons
bbox <- shp@bbox

# Get centroids
centroids <- coordinates(shp)
mo_shp <- shp
mo_shp$long_c <- centroids[,1]
mo_shp$lat_c <- centroids[,2]
mo_data <- mo_shp@data
mo_data <- mo_data %>% 
    left_join(districts, by = c("OKRUGS" = "English Name")) %>% 
    left_join(areas, by = c("RAION" = "English"))
mo_shp$district <- mo_data$`Russian Name`
mo_shp$area <- mo_data$Russian
mo_shp$district <- factor(mo_shp$district)
mo_shp$area <- factor(mo_shp$area)
head(mo_shp)
```

```{r}
okrugs_pal <- brewer.pal(12, "Paired") # Define a nice color palette with RColorBrewer
plot(mo_shp, border = "#57595b", col = okrugs_pal[as.numeric(mo_shp$district)], main = "Округи и Районы")
points(centroids, col = "black", pch = 20, cex = 0.3)
legend("right", title = "Округ", legend = sort(unique(mo_shp$district)),  fill = okrugs_pal, border = "#57595b", cex=0.8)
```

```{r}
# These plots can be interchanged between Russian and English names. I hate transliterated Russian, so almost all plots that I do are in Russian, but all that needs to be done is to switch the column used in the dataset; District corresponds with OKRUGS and area with RAION
spplot(mo_shp, c("district"), main = "Округи и районы Москвы", col.regions=okrugs_pal)
```

I will add some summary characteristics of houses in each area
```{r}
head(train)

area_frame <- train %>%
    group_by(sub_area) %>%
    summarize(MedYr = median(build_year, na.rm = TRUE),
              AvgSalePrice = mean(price_doc),
              AvgSqFt = mean(full_sq))
area_frame
head(area_frame)
head(mo_shp)
mo_data <- mo_data %>% left_join(area_frame, by = c("RAION" = "sub_area"))
area_frame
head(mo_data)
mo_shp@data <- mo_data
mo_shp$build_year <- mo_data$MedYr
mo_shp$AvgSalePrice <- mo_data$AvgSalePrice
mo_shp$AvgSqFt
head(mo_shp)
ncolors <- 15
pal3 <- colorRampPalette(c("grey90", "red"))(ncolors)
spplot(mo_shp, c("AvgSqFt"), main = "Transaction prices", col.regions = pal3,
       sub = "", cuts = ncolors-1, col = "transparent")
```

